---
title: "Final_Project_Script"
output: html_document
author: "<u>STAT400 Group #5:</u> Eugene Domrachev, Asher Eisen, Benjamin Cai, Urvi Chowdhury, Sebastian Gutierrez, Karn Randhawa, Keonwoo Kim, Daniel Zafman, Sahil Sinha, Ashish Dasu, Edward Du, Alex Mitin Ben Snyder, Brian Ndofor"
---

<h2> OBJECTIVE: </h2>

<h3><i>How does the population density and economic status of a state affect the COVID recovery rate?</i></h3>


<h2> HYPOTHESIS: </h2>

1. We expect that the higher the population density in a region, the lower the recovery rate will be. 
2. We expect that the higher the economic status is in a region, the higher the recovery rate will be.



```{r open csvs }
library(dplyr)
# install.packages("ggplot2") if you are missing it
library(ggplot2)



# rework the data to an easily plottable format

transpose_and_rename <- function(df){
   
   # transpose the data, so that making visuals is easier
  df <- as.data.frame(t(df))
  
  # make the states the column names
  names(df) <- df %>% slice(1) %>% unlist()
  
  # slice off the first row, where the state names were (they are labels, and should not be processed with other data)
  df <- df %>% slice(-1)
  
  # convert columns to numeric data, instead of factor
  df$RI <- as.numeric(df$RI)
  df$MD <- as.numeric(df$MD)
  df$VA <- as.numeric(df$VA)
  df$DC <- as.numeric(df$DC)
  
  return(df)
  
}
 
 

active_cases <- read.csv("../data/selected/active.csv")
active_cases <- transpose_and_rename(active_cases)

recovery_total <- read.csv("../data/selected/cumulative_recoved_total.csv")
recovery_total <- transpose_and_rename(recovery_total)

cumulative_recovery_rate <- read.csv("../data/selected/cumulative_recovery_rate.csv")
cumulative_recovery_rate <- transpose_and_rename(cumulative_recovery_rate)

recovery_proportion <- read.csv("../data/selected/daily_recovery_proportion.csv")
recovery_proportion <- transpose_and_rename(recovery_proportion)

```


``` {r check NA}

# Ensured no possibility of NA data points influencing further data 
# collection/manipulation.

active_cases <- active_cases[complete.cases(active_cases), ]
recovery_total <- recovery_total[complete.cases(recovery_total), ]
cumulative_recovery_rate <- cumulative_recovery_rate[complete.cases(cumulative_recovery_rate), ]
recovery_proportion <- recovery_proportion[complete.cases(recovery_proportion), ]


```


``` {r return/display dfs}

# Displayed all dataframes

head(active_cases)
head(recovery_total)
head(cumulative_recovery_rate)
head(recovery_proportion)

```

``` {r cumulative recovery rate}

factored_dates <- as.factor(rownames(recovery_total))
colors <- c(MD="#5999E3", VA="#E386C9", RI="#E3B159", DC="#77E364")


# ================ cumulative recovery rate ================

line_size <- 1.2

ggplot(data=recovery_proportion, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=cumulative_recovery_rate, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=cumulative_recovery_rate, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +
  geom_line(data=cumulative_recovery_rate, aes(x=factored_dates, y=DC, group=4, color="DC"), size=line_size) + 
  
  # scales, titles, labels
  #scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Cumulative Recovery Rate") + 
  ylab("Proportion Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5"),
         aspect.ratio=6/10
        )




# ================ cumulative recovery rate, without DC ================

ggplot(data=recovery_proportion, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=cumulative_recovery_rate, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=cumulative_recovery_rate, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +

  # scales, titles, labels
  #scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_y_continuous(limits = c(0, 0.18)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Cumulative Recovery Rate") + 
  ylab("Proportion Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5"),
         aspect.ratio=2/10
        ) 



  
```

``` {r daily recovery proportion}


# ================ proportional recovery rate ================

line_size <- 1
ggplot(data=recovery_proportion, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=recovery_proportion, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=recovery_proportion, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +
  geom_line(data=recovery_proportion, aes(x=factored_dates, y=DC, group=4, color="DC"), size=line_size) + 
  
  # scales, titles, labels
  #scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_y_continuous(limits = c(0, 0.065)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Daily Recovery Proportion") + 
  ylab("Proportion Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5"),
         aspect.ratio=6/10
        ) 








# ================ daily recovery proportion in non-DC states ================

line_size <- 0.6
ggplot(data=recovery_proportion, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=recovery_proportion, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=recovery_proportion, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +
  #geom_line(data=recovery_proportion, aes(x=factored_dates, y=DC, group=4, color="DC"), size=line_size) + 
  
  # scales, titles, labels
  #scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_y_continuous(limits = c(0, 0.005)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Daily Recovery Proportion") + 
  ylab("Proprtion Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5"),
        aspect.ratio=3/10
        ) 

```



```{r cumulative recovery total}

# ================ cumulative recovery totals ================

line_size <- 1.2
ggplot(data=recovery_total, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=DC, group=4, color="DC"), size=line_size) + 
  
  # scales, titles, labels
  scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Cumulative Recovery Total") + 
  ylab("Total Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5")
        ) 
  




```
```{r summary statistics}

# ================ cumulative recovery totals ================

line_size <- 1.2
ggplot(data=recovery_total, aes(x=factored_dates, y=RI, group=1, color="RI")) +
  
  #plotting the line data
  geom_line(size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=VA, group=2, color="VA"), size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=MD, group=3, color="MD"), size=line_size) +
  geom_line(data=recovery_total, aes(x=factored_dates, y=DC, group=4, color="DC"), size=line_size) + 
  
  # scales, titles, labels
  scale_y_continuous(breaks=seq(0, 25500, 5000), limits = c(0, 25500)) +
  scale_x_discrete(breaks=c("X04.12.2020", "X05.01.2020", "X06.01.2020", 
                          "X07.01.2020", "X08.01.2020", "X09.01.2020", 
                          "X10.01.2020", "X11.01.2020", "X12.01.2020")
                 , labels=c("April","May","June",
                            "July", "August", "September", 
                            "October", "November", "December")) +
  ggtitle("Cumulative Recovery Total") + 
  ylab("Total Recovered") + 
  xlab("Time") +
  
  scale_colour_manual(values=colors)+
  
  # general theme
  theme(
        text=element_text(size=12,family="serif"),
        plot.title = element_text(size=20, hjust = 0.5, vjust=3),
        axis.title.x = element_text(size=16, vjust=-2, hjust = 0.47),
        axis.title.y = element_text(size=14, vjust=5),
        plot.margin = unit(c(0.75, 0.5, 0.75, 0.75), "cm"),
        legend.position="right",
        legend.title = element_blank(),
        legend.key.size = unit(1, "cm"),
        legend.key.width = unit(1,"cm"),
        legend.background = element_rect(fill="transparent"),
        legend.box.background = element_rect(fill="transparent"),
        # change background of whole thing
        plot.background = element_rect(fill = "#f9faf5"),
        # change background of inner panel
        panel.background = element_rect(fill = "#dfe7f5")
        ) 
  




```

<h2> CONCLUSION & FUTURE DIRECTION </h2>

Daily and cumulative recovery proportions were consistent with our hypothesis about the relationship between population density and recovery rate for VA, MD, RI. We did not observe a relationship between economic status and recovery rate. 

Missing data and poor quality of data caused discrepancies and negated any expected predictions for DC. This problem was echoed for a number of states during the tenure of the project, among other problems:
<ul>
  <li> Some states were updated inconsistently (anywhere from 3-7 days in between updates), or stopped updating </li> 
  <li> Recovery totals were not available in some states </li> 
  <li> Fishy numbers (like the DC jump) </li> 
</ul>



